---
title: "Exercise_01"
output: html_notebook
---

The following table gives the volume, area, length and maximum and mean depths of some Scottish
lakes[1]. Create vectors, holding the lake’s name and all the parameters and build a dataframe
called scottish.lakes from the vectors
1 evaluate the highest and lowest volume and area lake
2 order the frame with respect to the area and determine the two largest area lakes
3 by summing up the areas occpupied by the lakes, determine the area of Scotland covered by water



```{r}
#EXERCISE 1

Loch <- c("Loch Ness", "Loch Lomond", "Loch Morar", "Loch Tay", "Loch Awe", "Loch Maree", "Loch Ericht", "Loch Lochy", "Loch Rannoch", "Loch Shiel", "Loch Katrine", "Loch Arkaig", "Loch Shin") 
Volume <- c(7.45, 2.6, 2.3, 1.6, 1.2, 1.09, 1.08, 1.07, 0.97, 0.79, 0.77, 0.75, 0.35) 
Area <- c(56, 71, 27, 26.4, 39, 28.6, 18.6, 16,19, 19.5, 12.4, 16, 22.5) 
Length <- c( 39,36, 18.8, 23, 41, 20, 23, 16, 15.7, 28, 12.9, 19.3, 27.8) 
Max_depth <- c(230, 190, 310, 150, 94, 114, 156, 162, 134, 128, 151, 109, 49) 
Mean_depth <- c(132, 37, 87,60.6, 32, 38, 57.6, 70, 51, 40, 43.4, 46.5, 15.5) 
scottish.lakes <- data.frame(Loch,Volume,Area,Length, Max_depth,Mean_depth )
```
```{r}
#Printing max and min values of volume and area
Vol_max <- max(scottish.lakes$Volume)
sprintf("Maximum volume: %f", Vol_max)

Vol_min <- min(scottish.lakes$Volume)
sprintf("Minimum volume: %f", Vol_min)

A_max <- max(scottish.lakes$Area)
sprintf("Maximum area: %f", A_max)

A_min <- min(scottish.lakes$Area)
sprintf("Minimum area: %f",A_min)


```

```{r}
#Sorting the dataframe by descending area values
scottish.lakes <- scottish.lakes[order(-scottish.lakes$Area), ]
scottish.lakes

```
```{r}
#Printing first 2 largest area lake names
print("largest lakes are:")
head(scottish.lakes$Loch, n = 2)

```
```{r}
#Summing elements of area column
Area_tot <- sum(scottish.lakes$Area)
sprintf("Total lake areas: %f", Area_tot)
```




The following CSV file
https://drive.google.com/file/d/13WkinYhJNSr_XCldb3wpZOI0fgjer7Iw/view contains data
on crude oil prices from 1861 to 2020, measured in US dollars per barrel [2].
1 Write R code that is able to read the file and import it in a data frame structure. (Hint: before
loading the file, open it with a text editor and check its structure).
2 produce a plot with the Oil price as a function of the year
3 which is the highest price in history ? When did it occur ?
3 plot the derivative of the curve, simply evaluated with the finite difference formula (forward derivative):


```{r}
#EXERCISE 2

#Extracting data from csv file
crude_oil_prices <- read.csv("crude-oil-prices.csv", header = TRUE)
crude_oil_prices
```
```{r}
#producting a plot with prices as function of the years 

plot(crude_oil_prices$Year, crude_oil_prices$Oil...Crude.prices.since.1861..current..., main = "Variation of crude oil prices", xlab = "Time [Year]", ylab = "Price [$]", xaxt = "n", pch = 19, cex= 0.7, col=3)

axis(side = 1, at = seq(1860, 2020, by = 10), las=2)

grid()

lines(crude_oil_prices$Year, crude_oil_prices$Oil...Crude.prices.since.1861..current..., col = "dark green")
```
```{r}

#Which is the highest price in history ? When did it occur ?

max_p <- max(crude_oil_prices$Oil...Crude.prices.since.1861..current...)
sprintf("Highest price of crude oil is: %f$", max_p)


max_y <- crude_oil_prices$Year[crude_oil_prices$Oil...Crude.prices.since.1861..current... == max_p]
sprintf("The maximum price occurs in: %i", max_y)
```
```{r}
#plot the derivative of the curve, simply evaluated with the finite difference formula


#Calculating a vector containing the derivatives (160 -1 elements)
prices <- crude_oil_prices$Oil...Crude.prices.since.1861..current...
l <- length(prices)
derivative <- prices[-1] - prices[-l]  #subtracting two tibbles 159X1: the first starts from second element, the second does not have the last element

#Plotting forward derivative (time goes from 1861 to 2020-1=2019)
plot(crude_oil_prices$Year[-l], derivative, main = "Derivative of crude oil prices", xlab = "Time [Year]", ylab = "Price per year [$ / year]", xaxt = "n", pch = 19,  cex= 0.7, col=3)

axis(side = 1, at = seq(1860, 2020, by = 10), las=2)

grid()

lines(crude_oil_prices$Year[-l], derivative, col = "dark green")
```


The following CSV file
https://drive.google.com/file/d/1U1hK5o_d3vl8twwGaRJNcDJ79O9I5zJ1/view?usp=sharing
contains data on the coal production (in TW · hour) for several countries in the World [3].
1 Write R code that is able to read the file and import it in a tibble [4] structure
2 count the number of countries available in the file and produce a barplot with the number of entries
for each country
for the following items select only the years ≥ 1970:
3 selecting only the year after 1970, determine the total integrated production for each country and
print the top 5 Countries with highest coal productions
4 for the 5 top Countries, create a plot of production as a function of time
5 generate a plot with the cumulative sum of the World’s coal production over the years




```{r}
#EXERCISE 3
#Write R code that is able to read the file and import it in a tibble structure:
library(tibble)


coal_production <- as_tibble(read.csv("coal-production-by-country.csv", header = TRUE))
coal_production



```
```{r}
#count the number of countries available in the file and produce a barplot with the number of entries for each country

countries <- coal_production$Entity[!duplicated(coal_production$Entity) ]  
sprintf("Number of countries in the dataframe is: %i", length(countries))
#countries

#Histogram
barplot(table(coal_production$Entity), ylab = "Coal production(TWh)")  #eventualmente da sistemare asse x

```
```{r}
#selecting only the year after 1970, determine the total integrated production for each country and print the top 5 Countries with highest coal productions
library(dplyr)

#Filtering by year 1970   
coal_production2 <- filter(coal_production, Year >= 1970)
int_production <- aggregate(coal_production2$Coal.production..TWh., by=list(coal_production2$Entity), FUN=sum)
int_production <- int_production[order(-int_production$x),]
print("First five countries for production are:")
int_production$Group.1[2:6]                          #from 2 to 6 because first elements is "world"

```


```{r}
#for the 5 top Countries, create a plot of production as a function of time
library(plotly)

top_prod <- filter(coal_production2, Entity == int_production$Group.1[2:6] ) #a tibble with top 5 countries information 

plot_ly(data = top_prod, x = ~Year, y=~ Coal.production..TWh., color = ~Entity, colors = "Set1", mode="lines+markers") %>% layout(title="Coal production top 5 countries")

```









```{r}
#EXERCISE 4
#filter the original tibble by selecting the following countries: Italy

vaccination_data <- as_tibble(read.csv("vaccinations-by-manufacturer.csv", header = TRUE))

vaccination_data_IT <- filter(vaccination_data, location == 'Italy')

vaccination_data_IT
```
```{r}
#plot the number of vaccines given as a function of time for the different vaccine manufacturer
library(plotly)

plot_ly(data = vaccination_data_IT, x = ~date, y=~total_vaccinations, color =~vaccine, colors = "Set1", mode="", opacity=1 ) %>% layout(title="Vaccines in Italy")
```



```{r}
#from the same tibble plot the total number of vaccines shot per day in Italy

tot_vac <- aggregate(vaccination_data_IT$total_vaccinations, list(vaccination_data_IT$date), sum )
tot_vac
plot(strptime(tot_vac$Group.1,"%Y-%m-%d"), tot_vac$x, col=4, pch = 3, cex= 0.5, xlab="Time [Dates]", ylab="Total Vaccinations", main="Total vaccination per day in Italy" )
grid()
```
```{r}
#do the same exercise for the following countries: Germany and United States of America

vaccination_data_USA <- filter(vaccination_data, location == 'United States')
vaccination_data_DE <- filter(vaccination_data, location == 'Germany')
vaccination_data_USA
vaccination_data_DE

plot_ly(data = vaccination_data_USA, x = ~date, y=~total_vaccinations, color =~vaccine, colors = "Set1", mode="lines+markers") %>% layout(title="Vaccines in USA")
plot_ly(data = vaccination_data_DE, x = ~date, y=~total_vaccinations, color =~vaccine, colors = "Set1", mode="lines+markers") %>% layout(title="Vaccines in Germany")


tot_vac_USA <- aggregate(vaccination_data_USA$total_vaccinations, list(vaccination_data_USA$date), sum )
plot(strptime(tot_vac_USA$Group.1, "%Y-%m-%d"), tot_vac_USA$x, col=4, pch = 3, cex= 0.5,xlab="Time [Dates]", ylab="Total Vaccinations", main="Total vaccination per day in USA" )
grid()


tot_vac_DE <- aggregate(vaccination_data_DE$total_vaccinations, list(vaccination_data_DE$date), sum )
plot(strptime(tot_vac_DE$Group.1,"%Y-%m-%d"), tot_vac_DE$x, col=4, pch = 3, cex= 0.5, xlab="Time [Dates]", ylab="Total Vaccinations", main="Total vaccination per day in Germany" )
grid()


```
```{r}
#selecting all the European countries in the tibble, plot the number of daily vaccinations per million as a function of date

vaccinations<- as_tibble(read.csv("vaccinations.csv", header=TRUE))
vaccinations_EUR <-  filter(vaccinations, iso_code == 'OWID_EUR')
vaccinations_EUR

plot(strptime(vaccinations_EUR$date, "%Y-%m-%d"), vaccinations_EUR$daily_vaccinations_per_million, col=4,  pch = 19, cex= 0.5, ylab="Daily vaccination per million", xlab="Time [Dates]", main="Vaccinations in European countries")
grid()

```
```{r}
#study the data structure and produce few relevant plots of your taste

plot(strptime(vaccinations_EUR$date,"%Y-%m-%d"), vaccinations_EUR$people_vaccinated, col="blue",type="b",  pch = 19, cex= 0.5, ylab='', xlab="Time [Dates]", main="Vaccinations in Europe")
points(strptime(vaccinations_EUR$date,"%Y-%m-%d"),vaccinations_EUR$total_vaccinations,col="green",type="b", pch = 19, cex= 0.5 )

legend(x="topleft", legend=c("people vaccinated", "total vaccination"),
       col=c("blue", "green"), lty=1:1, cex=1)
grid()


plot(strptime(vaccinations_EUR$date,"%Y-%m-%d"), vaccinations_EUR$daily_vaccinations_raw, col="blue",type="b",  pch = 19, cex= 0.5, ylab='', xlab="Time [Dates]", main="Vaccinations in Europe")
points(strptime(vaccinations_EUR$date,"%Y-%m-%d"),vaccinations_EUR$daily_vaccinations,col="green",type="b", pch = 19, cex= 0.5 )
grid()

legend(x="topleft", legend=c("Daily vaccination raw", "Daily vaccination"),
       col=c("blue", "green"), lty=1:1, cex=1)


plot(strptime(vaccinations_EUR$date,"%Y-%m-%d"), vaccinations_EUR$people_vaccinated_per_hundred, col="blue",type="b",  pch = 19, cex= 0.5, ylab='', xlab="Time [Dates]", main="Vaccinations in Europe")
points(strptime(vaccinations_EUR$date,"%Y-%m-%d"),vaccinations_EUR$people_fully_vaccinated_per_hundred,col="green",type="b", pch = 19, cex= 0.5 )
grid()

legend(x="topleft", legend=c("People vaccinated per hundred", "people fully vaccinated per hundred"),
       col=c("blue", "green"), lty=1:1, cex=1)


```













